Adaptative neighbouring methods applied to telephonic base stations

### 2024-08-23 article.tex
Retrait glossaire et imports inutiles

\documentclass[lettersize,journal,english]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{booktabs}
\usepackage{units}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]{hyperref}
\hypersetup{
 urlcolor=blue, linkcolor=black}
\usepackage[acronym]{glossaries}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

\global\long\def\BDsites{\textsf{BD\_sites}}

@@

\makeglossaries

\setglossarypreamble{All definitions are from \href{https://www.dictionary.com/}{Dictionnary.com}}

\newglossaryentry{base station}
{
    name=base station,
    description={A fixed transmitter that forms part of an otherwise mobile radio network}
}

\newglossaryentry{convHull}
{
    name=convex hull,
    description={The smallest convex set containing a given set}
}

\newacronym{arcep}{ARCEP}{French Regulatory Authority for Electronic Communications and Posts}
\newacronym{anfr}{ANFR}{French National Frequency Agency}
\newacronym{bs}{BS}{\gls{base station}}
\newacronym{dbscan}{DBScan}{Density-Based Spatial Clustering of Applications with Noise}
\newacronym{hdbscan}{HDBScan}{Hierarchical Density-Based Spatial Clustering of Applications with Noise}

@@

\begin{abstract}
    Nowadays, it is primordial for telecommunication companies to be able to determine if a user is moving. For that, it is necessary to understand the neighbouring relationships between the telephonic \acrlong{bs}s. This article aims at proposing and optimizing a method that adaptively detects these relationship between telephonic \acrlong{bs}s.

@@

    A cellular network, also know as a mobile network, is a type of telecommunications network that is dispersed over geographical areas known as cells, each of which is supplied by at least one fixed-location transceiver called \acrfull{bs}. These \acrshort{bs}s give the cell network coverage so that voice, data, and other kinds of content may be transmitted.

@@

    The field of neighbouring problems mainly revolves around the optimisation of the $k$ nearest neighbours algorithm ($k$-NN).
    As indicated by its name, this method consists in considering the closest $k$ points of each points as their neighbour.
    [insert KNN reference]

@@

    The default of the $k$-NN method is that it always find the same number of neighbours to each point, independantly to the reality of the geometry around each \acrshort{bs}.

@@

    Here two \acrshort{bs}s are considered as neighbours if their converage area overlap. Therefore, these generic methods aren't perfectly adapted to this application as these coverage areas are submitted to specific limitations like limitted radius and wave interferences.

@@

    Therefore, the outcome of the method presented is a neighbouring graph, cf. Figure~\ref{fig:neigh_graph}, where each node is a \acrshort{bs} and each edge represent a neighbouring relationship between these two \acrshort{bs}s.
    \begin{figure}
        \centering
        \includegraphics[width=3.5in]{images/illus_graphs/neighbouring_graph.png}
        \caption{Illustration of a neighbouring graph.}
        \label{fig:neigh_graph}
    \end{figure}

    A key finding of this article is to locally take into account the density of \acrshort{bs} in the neighbouring methodology.

@@

    \subsection{Main database}
        In this study, this database \cite{main_database} that contains all the information needed for the application of the method is being used. It is from \acrfull{arcep}, which is an independent French administrative authority responsible for regulating electronic and postal communications and press distribution.

@@

        It consists of a substantial collection of $108\,838$ entries, encompassing $29$ distinct features of a size of $\unit[16.7]{Mo}$.
        Each row of the database corresponds to detailed information regarding a specific BS. For the scope of this study, our analysis is focusing on the fields outlined in Table~\ref{table:data_columns}.
        \begin{table}
            \centering
            \caption{Description of the main database columns used in the experiments.}
            \label{table:data_columns}
            \begin{tabular}{ll}
                \toprule
                \textbf{Column} & \textbf{Description} \\
                \cmidrule(lr){1-2}
                \textsl{id\_station\_anfr} & \acrshort{anfr} \acrshort{bs} ID \\ 
                \textsl{nom\_op} & Name of the provider \\
                \textsl{$x$, $y$, latitude, longitude} & Base station coordinates \\ 
                \textsl{nom\_reg, nom\_dep, nom\_com} & Additional location information \\  
                \textsl{site\_2g, 3g, 4g, 5g} & Technology used by the \acrshort{bs} \\ 
                \bottomrule
            \end{tabular}
        \end{table}

@@

        Finally, informations about the technologies used in every site is provided. Indeed, a site can use multiple technologies.
        In Figure~\ref{fig:data_evolution} the evolution of each technology from the first updates of the data can be found.
        \begin{figure}
            \centering
            \includegraphics[width=3.5in]{images/data_analysis/technos-evolution.png}
            \caption{Evolution by quarter year and by technology of the number of \acrshort{bs}.}
            \label{fig:data_evolution}
        \end{figure}
        Then, in Figure~\ref{fig:data_technos} a chart showing the number of times each technology is present for each provider can be found.
        \begin{figure}
            \centering
            \includegraphics[width=3.5in]{images/data_analysis/with_techno.png}
            \caption{Breakdown, by operator, of the number of sites containing a given technology.}
            \label{fig:data_technos}
        \end{figure}
        We can see, for example, that \emph{Free Mobile} doesn't have any \acrshort{bs} with $2$G technology. Additionally, you can also find these values in Table~\ref{table:techno_numbers}.
        \begin{table}
            \centering
            \caption{Number of \acrshort{bs} equipped, at least, with $x$G, by provider.}
            \label{table:techno_numbers}
            \begin{tabular}{cccccc}
                \toprule
                \textbf{Technology} & \textbf{Bouygues} & \textbf{Free} & \textbf{Orange} & \textbf{SFR} & \textbf{Total} \\
                \cmidrule(lr){1-6}
                \textbf{2G} & 21710 & 0 & 18835 & 20093 & 60638 \\
                \textbf{3G} & 26294 & 25934 & 29748 & 25777 & 107753 \\
                \textbf{4G} & 26231 & 25847 & 30440 & 25881 & 108399 \\
                \textbf{5G} & 11271 & 18607 & 8794 & 10968 & 49640 \\
                \textbf{Total} & 26331 & 25949 & 30540 & 26018 & 108838 \\
                \bottomrule
            \end{tabular}
        \end{table}

@@

        Firstly, only one provider is considered because a \acrshort{bs} from provider $x$ can't be considered as a neighbour of \acrshort{bs} from provider $y$.
        Consequently, we have chosen \emph{Orange} because it is the one that has the largest number of \acrshort{bs}.

        Secondly, for the same reasons as before, only one technology is considered. Thus, all \acrshort{bs} using, at least, 4G are taken into account.

@@

        \subsubsection{Delaunay triangulation}
            A Delaunay triangulation \cite{art_delaunay} of a set of points in the plane subdivides their \gls{convHull} into triangles whose circumcircles do not contain any of the points.

@@

    \subsection{Finding real neighbours}
        For the purpose of our experiments, i.e. finding \acrshort{bs} neighbours, the previously presented methods are not sufficient to find real neighbours.
        As a consequence, some neighbouring connexions needs to be suppressed \cite{patent_neighs}.
        \subsubsection{Distance criterion}
            A \acrshort{bs} does not have an infinite coverage area.
            As a consequence, each edge in $U$ that is longer than a certain value, called \emph{max\_distance}, is suppressed.

        \subsubsection{Angle criterion}
            Let $p$, $q$, $r$ be three \acrshort{bs} of $P$, where $q$ and $r$ are $p$'s potential neighbours. As said before, coverage areas of \acrshort{bs} are limited by physics, this also implies that $r$ can be \og hidden\fg{} from $p$ by $q$ because $q$ is between $p$ and $r$ [trouver qqch pour prouver ce que je dis].

@@

            It consists in dividing the space around each \acrshort{bs} into quadrants and keeping only a certain number of potential neighbours in each of the quadrants, cf. Figure~\ref{fig:crit_qua}.
            The number of quadrants has been fixed to six \cite{art_del_paq}.

    \subsection{Density-based clustering methods}
        \subsubsection{\acrshort{dbscan}}
            \acrshort{dbscan} consists in detecting clusters based on the density of points, using two parameters:
            \begin{itemize}
                \item $\varepsilon$: A distance threshold under which two points are considered close enough to be considered neighbours;   
                \item $n_{\text{min}}$: The shortest number of neighbours a points needs to have to be considered as a core point of a cluster.
            \end{itemize}

@@

        \subsubsection{\acrshort{hdbscan}}
            \acrshort{hdbscan} \cite{10.1007/978-3-642-37456-2_14} is a variation of the \acrshort{dbscan} method.
            It consists in selecting the best clusters after testing their persistence by making the $\varepsilon$ parameter vary. It's main particularity is that it allows to detect clusters of different densities.

@@

\section{Contributions\label{sec:contrib}}
    \noindent In section~\ref{sec:overview}, we have seen that we can find the neighbours of \acrshort{bs}s by combining a potential neighbouring graph and filtration criteria. This methodology gives pretty good results \cite{art_del_paq}.

    However, in certain cases like in a city, the filtration criteria aren't precise enough. We, therefore, need to find some ways to measure \acrshort{bs}s' density.

    \subsection{Measurement of \acrshort{bs} density}
        Places with a high population also have a high density of \acrshort{bs}s. In fact, each \acrshort{bs} can only support a certain load of users.

        Therefore, let's assume that finding zones with a high density of \acrshort{bs} is equivalent to finding the \acrshort{bs} situated in \og cities\fg{}.

        \subsubsection{\acrshort{dbscan}}
            The most intuitive method to use when trying to locate zones with high a high density of point is \acrshort{dbscan}. In fact, by applying the methods with the right parameters, these zones are detected in separate clusters and the rest is considered as noise. Two categories are therefore created: city (all the clusters) and countryside (the noise).

        \subsubsection{\acrshort{hdbscan}}
            We can use \acrshort{hdbscan} the same way. However, the previously mentionned probability associated with each point belonging to a cluster can be used to separate the center of the city (where the probability is very high) from the rest of the city.

        \subsubsection{$3$-NN mean distance}
            For each \acrshort{bs} $p\in P$, the mean distance $\gamma_p$, in $\unit{km}$, to its three nearest neighbours is calculated. According to the value of $\gamma_p$, a \acrshort{bs} is classified into a different category of density. In fact, a low $\gamma_p$ means that the surrounding zone is very dense in \acrshort{bs}. On the contrary, a high value translate to a low density area (countryside).

            For this method to work properly, only \acrshort{bs} from the same provider have to be kept into the calculation process.

@@

        \subsubsection{Distance and angle criterion}
            As seen in section~\ref{sec:overview}, after having found potential neighbours for each \acrshort{bs}, this connexions needs to be filtrated. However, this applies the same distance and angle threshold for all \acrshort{bs}s, without taking in account the difference of density.

@@

            For both distance and angle, the more the \acrshort{bs} is in a dense category, the lower this threshold is.

@@

\section{Results\label{sec:res}}
    \subsection{Measurement of \acrshort{bs} density}

@@

        In Figure~\ref{fig:final_neighs} you have an illustration of a portion of the results. To clarify, the colour of the points, which represent the \acrshort{bs}s, is in function of the $\gamma_p$ value.

@@

\printglossary[type=\acronymtype]
\printglossary

>>>>

\documentclass[lettersize,journal,english]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage{algorithmic}
% \usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{booktabs}
\usepackage{units}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=true]{hyperref}
\hypersetup{
 urlcolor=blue, linkcolor=black}

@@

@@

\begin{abstract}
    Nowadays, it is primordial for telecommunication companies to be able to determine if a user is moving. For that, it is necessary to understand the neighbouring relationships between the telephonic base stations. This article aims at proposing and optimizing a method that adaptively detects these relationship between telephonic base stations.

@@

    A cellular network, also know as a mobile network, is a type of telecommunications network that is dispersed over geographical areas known as cells, each of which is supplied by at least one fixed-location transceiver called base station (BS). These BSs give the cell network coverage so that voice, data, and other kinds of content may be transmitted.

@@

    The field of neighbouring problems mainly revolves around the optimisation of the $k$ Nearest Neighbours ($k$-NN) algorithm.
    As indicated by its name, this method consists in considering the closest $k$ points of each points as their neighbour.
    [insert KNN reference]

@@

    The default of the $k$-NN method is that it always find the same number of neighbours to each point, independantly to the reality of the geometry around each BS.

@@

    Here two BSs are considered as neighbours if their converage area overlap. Therefore, these generic methods aren't perfectly adapted to this application as these coverage areas are submitted to specific limitations like limitted radius and wave interferences.

@@

    Therefore, the outcome of the method presented is a neighbouring graph, cf. Figure~\ref{fig:neigh_graph}, where each node is a BS and each edge represent a neighbouring relationship between these two BSs.
    \begin{figure}
        \centering
        \includegraphics[width=3.5in]{images/illus_graphs/neighbouring_graph.png}
        \caption{Illustration of a neighbouring graph.}
        \label{fig:neigh_graph}
    \end{figure}

    A key finding of this article is to locally take into account the density of BS in the neighbouring methodology.

@@

    \subsection{Main database}
        In this study, this database \cite{main_database} that contains all the information needed for the application of the method is being used. It is from the French Regulatory Authority for Electronic Communications and Posts (ARCEP), which is an independent French administrative authority responsible for regulating electronic and postal communications and press distribution.

@@

        It consists of a substantial collection of $108\,838$ entries, encompassing $29$ distinct features of a size of $\unit[16.7]{Mo}$.
        Each row of the database corresponds to detailed information regarding a specific BS. For the scope of this study, our analysis is focusing on the fields outlined in Table~\ref{table:data_columns}.
        \begin{table}
            \centering
            \caption{Description of the main database columns used in the experiments.}
            \label{table:data_columns}
            \begin{tabular}{ll}
                \toprule
                \textbf{Column} & \textbf{Description} \\
                \cmidrule(lr){1-2}
                \textsl{id\_station\_anfr} & ANFR BS ID \\ 
                \textsl{nom\_op} & Name of the provider \\
                \textsl{$x$, $y$, latitude, longitude} & Base station coordinates \\ 
                \textsl{nom\_reg, nom\_dep, nom\_com} & Additional location information \\  
                \textsl{site\_2g, 3g, 4g, 5g} & Technology used by the BS \\ 
                \bottomrule
            \end{tabular}
        \end{table}

@@

        Finally, informations about the technologies used in every site is provided. Indeed, a site can use multiple technologies.
        In Figure~\ref{fig:data_evolution} the evolution of each technology from the first updates of the data can be found.
        \begin{figure}
            \centering
            \includegraphics[width=3.5in]{images/data_analysis/technos-evolution.png}
            \caption{Evolution by quarter year and by technology of the number of BS.}
            \label{fig:data_evolution}
        \end{figure}
        Then, in Figure~\ref{fig:data_technos} a chart showing the number of times each technology is present for each provider can be found.
        \begin{figure}
            \centering
            \includegraphics[width=3.5in]{images/data_analysis/with_techno.png}
            \caption{Breakdown, by operator, of the number of sites containing a given technology.}
            \label{fig:data_technos}
        \end{figure}
        We can see, for example, that \emph{Free Mobile} doesn't have any BS with $2$G technology. Additionally, you can also find these values in Table~\ref{table:techno_numbers}.
        \begin{table}
            \centering
            \caption{Number of BS equipped, at least, with $x$G, by provider.}
            \label{table:techno_numbers}
            \begin{tabular}{cccccc}
                \toprule
                \textbf{Technology} & \textbf{Bouygues} & \textbf{Free} & \textbf{Orange} & \textbf{SFR} & \textbf{Total} \\
                \cmidrule(lr){1-6}
                \textbf{2G} & 21710 & 0 & 18835 & 20093 & 60638 \\
                \textbf{3G} & 26294 & 25934 & 29748 & 25777 & 107753 \\
                \textbf{4G} & 26231 & 25847 & 30440 & 25881 & 108399 \\
                \textbf{5G} & 11271 & 18607 & 8794 & 10968 & 49640 \\
                \textbf{Total} & 26331 & 25949 & 30540 & 26018 & 108838 \\
                \bottomrule
            \end{tabular}
        \end{table}

@@

        Firstly, only one provider is considered because a BS from provider $x$ can't be considered as a neighbour of BS from provider $y$.
        Consequently, we have chosen \emph{Orange} because it is the one that has the largest number of BS.

        Secondly, for the same reasons as before, only one technology is considered. Thus, all BS using, at least, 4G are taken into account.

@@

        \subsubsection{Delaunay triangulation}
            A Delaunay triangulation \cite{art_delaunay} of a set of points in the plane subdivides their convex hull into triangles whose circumcircles do not contain any of the points.

@@

    \subsection{Finding real neighbours}
        For the purpose of our experiments, i.e. finding BS neighbours, the previously presented methods are not sufficient to find real neighbours.
        As a consequence, some neighbouring connexions needs to be suppressed \cite{patent_neighs}.
        \subsubsection{Distance criterion}
            A BS does not have an infinite coverage area.
            As a consequence, each edge in $U$ that is longer than a certain value, called \emph{max\_distance}, is suppressed.

        \subsubsection{Angle criterion}
            Let $p$, $q$, $r$ be three BS of $P$, where $q$ and $r$ are $p$'s potential neighbours. As said before, coverage areas of BS are limited by physics, this also implies that $r$ can be \og hidden\fg{} from $p$ by $q$ because $q$ is between $p$ and $r$ [trouver qqch pour prouver ce que je dis].

@@

            It consists in dividing the space around each BS into quadrants and keeping only a certain number of potential neighbours in each of the quadrants, cf. Figure~\ref{fig:crit_qua}.
            The number of quadrants has been fixed to six \cite{art_del_paq}.

    \subsection{Density-based clustering methods}
        \subsubsection{DBScan}
            DBScan consists in detecting clusters based on the density of points, using two parameters:
            \begin{itemize}
                \item $\varepsilon$: A distance threshold under which two points are considered close enough to be considered neighbours;   
                \item $n_{\text{min}}$: The shortest number of neighbours a points needs to have to be considered as a core point of a cluster.
            \end{itemize}

@@

        \subsubsection{HDBScan}
            HDBScan \cite{10.1007/978-3-642-37456-2_14} is a variation of the DBScan method.
            It consists in selecting the best clusters after testing their persistence by making the $\varepsilon$ parameter vary. It's main particularity is that it allows to detect clusters of different densities.

@@

\section{Contributions\label{sec:contrib}}
    \noindent In section~\ref{sec:overview}, we have seen that we can find the neighbours of BSs by combining a potential neighbouring graph and filtration criteria. This methodology gives pretty good results \cite{art_del_paq}.

    However, in certain cases like in a city, the filtration criteria aren't precise enough. We, therefore, need to find some ways to measure BSs' density.

    \subsection{Measurement of BS density}
        Places with a high population also have a high density of BSs. In fact, each BS can only support a certain load of users.

        Therefore, let's assume that finding zones with a high density of BS is equivalent to finding the BS situated in \og cities\fg{}.

        \subsubsection{DBScan}
            The most intuitive method to use when trying to locate zones with high a high density of point is DBScan. In fact, by applying the methods with the right parameters, these zones are detected in separate clusters and the rest is considered as noise. Two categories are therefore created: city (all the clusters) and countryside (the noise).

        \subsubsection{HDBScan}
            We can use HDBScan the same way. However, the previously mentionned probability associated with each point belonging to a cluster can be used to separate the center of the city (where the probability is very high) from the rest of the city.

        \subsubsection{$3$-NN mean distance}
            For each BS $p\in P$, the mean distance $\gamma_p$, in $\unit{km}$, to its three nearest neighbours is calculated. According to the value of $\gamma_p$, a BS is classified into a different category of density. In fact, a low $\gamma_p$ means that the surrounding zone is very dense in BS. On the contrary, a high value translate to a low density area (countryside).

            For this method to work properly, only BS from the same provider have to be kept into the calculation process.

@@

        \subsubsection{Distance and angle criterion}
            As seen in section~\ref{sec:overview}, after having found potential neighbours for each BS, this connexions needs to be filtrated. However, this applies the same distance and angle threshold for all BSs, without taking in account the difference of density.

@@

            For both distance and angle, the more the BS is in a dense category, the lower this threshold is.

@@

\section{Results\label{sec:res}}
    \subsection{Measurement of BS density}

@@

        In Figure~\ref{fig:final_neighs} you have an illustration of a portion of the results. To clarify, the colour of the points, which represent the BSs, is in function of the $\gamma_p$ value.

@@
